---
title: "Midterm Proposal"
author: "Basmah"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#put all packages here
pacman::p_load(gtsummary,tidyverse,janitor,WDI)
```

## Primary Aim

My app will let the user explore health care insurance coverage and associate it to various health outcomes.
The users will be able to explore data series using an interactive R shiny app.

The app will provide an interface to the World data bank. 
I will pre-select ten or twenty series for the users to select.
The series will include measures of health insurance, health care accessibility, health outcomes and demographic information. 
The users can select data series, graphing methods and statistical analysis.

This app will help inform policy makers and tell them how health insurance and health care availability is related to health outcomes.

## Code Sample, Proof-of-concept


### Access the Data

The data will come from the World bank. 
I downloaded data to csv file, but my app will use the `WDI` app to download data as the users request it.

Below is an example of the outcome life expectancy .


```{r}
le <- read_csv("data-raw/life-expectancy.csv") %>% 
  rename(country_name=Entity, country_code=Code)
le %>% head
```
This is an example of health care expenditure. Health care expenditure is something that policy makers can control. Health policy makers will be able to measure the affect of expenditure an outcomes such as life expectancy.


```{r}
ce <- read_csv("data-raw/datac.csv") %>% 
  clean_names() %>% 
pivot_longer(starts_with("x"),names_to = "year", values_to = "expenditure") %>% 
  mutate(Year = as.numeric(str_extract(year,"(\\d{4})"))) %>% 
  mutate(expenditure = as.numeric(expenditure)) %>% 
  select(country_name, country_code, Year, expenditure) %>% 
  na.omit()
ce  %>% head
```
The app will also provide data series that can be used in the statistical models to adjust for national level characteristics. 
In this data set we show Adjusted net national income. 

```{r}
adj <- read_csv("data-raw/4bd3a9ba-b34d-4a2d-b9e3-9ff60a60b55e_Data.csv") %>%
  clean_names() %>% 
  pivot_longer(starts_with("x"),names_to = "year", values_to = "adjusted") %>% 
  mutate(Year = as.numeric(str_extract(year,"(\\d{4})"))) %>% 
  filter(series_code == "SH.XPD.EHEX.PP.CD") %>% 
  arrange(country_name, Year) %>% 
  select(country_name, country_code, Year, adjusted, everything()) %>% 
  mutate(adjusted = as.numeric(adjusted)) %>% 
  select(country_name, country_code, Year, adjusted) %>% 
  na.omit()
adj 
```
The users will be able to combine measures to create and analysis data set.

```{r}
d <- ce %>% 
  inner_join(adj, by = c('country_code','Year','country_name')) %>% 
  inner_join(le, by = c('country_code','Year','country_name')) %>%
  filter(!is.na(adjusted), !is.na(expenditure))
d %>% head(10)
```

The users will be able to select from data summary options.

```{r}
d %>% 
  select(-country_code, -Year) %>% 
  gtsummary::tbl_summary(by = c("country_name"))
```
The data can be saved to the package if needed.

```{r}
#usethis::use_data(adj,ce,d,overwrite = TRUE)
```

## Challenges 

1. Finding the data series.
2. Design the data plots.
3. Design the statistical analysis.
4. Learn R shiny.
5. Connect R shiny to the data source.


## To do list

1. Find the data series. (October 15)
2. Design the plots and data explorations. (October 22)
3. Create R shiny interface to the data. (October 29)
4. Create R shiny users interface.(November 5)
5. Write the R shiny server cod.(November 12)
6. Deploy the R shiny app.(November 20)


## install package

```{r}
remotes::install_github("Basmahsal/alharbi")
```






