---
title: "Getting data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(gtsummary,tidyverse,janitor,WDI)
```

## Chosen Indicators/Series

SH.UHC.NOP2.ZG
:  increase in poverty gap at $3.20 ($ 2011 PPP) poverty line due to out-of-pocket health care expenditure (% of poverty line)

SP.DYN.LE00.IN
:    

SH.XPD.OOPC.CH.ZS
:  Out of pocket expenditure

SH.XPD.PVTD.PP.CD
:  Domestic private health expenditure (per capite)

SH.XPD.EHEX.PP.CD
:  External health expenditure 




```{r}
series_choices <- c("SH.UHC.NOP2.ZG", 
                    "SH.XPD.CHEX.PP.CD" ,
                    "NY.ADJ.NNTY.KD.ZG" ,
                    "SP.DYN.LE00.IN" ,
                    "SH.XPD.EHEX.PP.CD",
                    "SH.XPD.PVTD.PP.CD",
                    "SH.XPD.GHED.PP.CD",
                    "SH.XPD.OOPC.CH.ZS")
```

```{r}
outcomes <- c("SH.UHC.NOP2.ZG", 
                    "SH.XPD.CHEX.PP.CD" ,
                    "SH.XPD.EHEX.PP.CD",
                    "SH.XPD.OOPC.CH.ZS")
```


```{r}
predicters <- c("NY.ADJ.NNTY.KD.ZG" ,
                    "SP.DYN.LE00.IN" ,
                    "SH.XPD.PVTD.PP.CD",
                    "SH.XPD.GHED.PP.CD")
```




```{r}
dd <- WDI( indicator = series_choices)
```

```{r}
usethis::use_data(dd, overwrite = TRUE)
```



```{r}
dd %>% 
  count(country)
```
```{r}
dd %>% 
  filter(country == 'Saudi Arabia')
```


```{r}
d <- dd %>%
  na.omit() 

d %>% head
```


```{r}
d1 <- read_csv("data-raw/59cc850e-8964-4231-bd2d-00676a3f3777_Data.csv")
d1 %>% head
```
```{r}
dd1 <- WDI(indicator = "SH.UHC.NOP2.ZG") 
```

```{r}
dd1 %>% head
```

```{r}
dd1 %>% 
  filter(!is.na(SH.UHC.NOP2.ZG )) %>% 
  group_by(country) %>% 
  summarize(mean = mean(SH.UHC.NOP2.ZG,na.rm = TRUE),.groups = 'drop') %>% 
  arrange(desc(mean))
```

## Extrnal health expenditure 
Series code
:   SH.XPD.EHEX.PP.CD

```{r}
ex <- WDI::WDI(indicator = "SH.XPD.EHEX.PP.CD")
ex %>% head
```

```{r}
usethis::use_data(ex, overwrite = TRUE)

```
```{r}
ex %>% 
  count(country, year) %>% 
  arrange(desc(n))
```
```{r}
ex %>% 
  filter(!is.na(SH.XPD.EHEX.PP.CD)) %>% 
  group_by(country) %>% 
  summarize(mean = mean(SH.XPD.EHEX.PP.CD,na.rm = TRUE),.groups = 'drop') %>% 
  arrange(desc(mean))
```

## Out of pocket expenditure
Series Code
:     SH.XPD.OOPC.CH.ZS
```{r}
OUT <- WDI::WDI(indicator = "SH.XPD.OOPC.CH.ZS" )
OUT %>% head
```
```{r}
usethis::use_data(OUT, overwrite = TRUE)

```

```{r}
OUT %>% 
  count(country, year) %>% 
  arrange(desc(n))
```

```{r}
OUT  %>% 
  filter(!is.na(SH.XPD.OOPC.CH.ZS)) %>% 
  group_by(country) %>% 
  summarise(mean=mean(SH.XPD.OOPC.CH.ZS, na.rm=TRUE), .groups="drop") %>% 
  arrange(desc(mean))
```




## Domestic private health expenditure (per capite)
Series Code
:    SH.XPD.PVTD.PP.CD
```{r}
p <- WDI::WDI(indicator = "SH.XPD.PVTD.PP.CD")

```

```{r}
usethis::use_data(p, overwrite = TRUE)
```
```{r}
p %>% 
  count(country, year) %>% 
  arrange(desc(n))
```

```{r}
p %>% 
  filter(!is.na(SH.XPD.PVTD.PP.CD)) %>% 
  group_by(country) %>% 
  summarise(mean=mean(SH.XPD.PVTD.PP.CD, na.rm=TRUE), .groups="drop") %>% 
  arrange(desc(mean))
  
```




## Life Expectancy at Birth(years)

Series code
:  SP.DYN.LIFE.MF

```{r}

WDIsearch('life expectancy')
```

```{r}
WDIsearch(string = 'Health expenditure',field = 'description')
```

```{r}
le <- WDI::WDI(indicator = "SP.DYN.LE00.IN")
le %>% head
```

```{r}
usethis::use_data(le, overwrite = TRUE)
```

```{r}
le %>% 
  count(country, year) %>% 
  arrange(desc(n))

```


## Domestic general govermnent health expenditure (per capite)
Series Code
:    SH.XPD.GHED.PP.CD
```{r}
dg <- WDI::WDI(indicator = "SH.XPD.GHED.PP.CD")
dg %>% head
```
```{r}
usethis::use_data(dg, overwrite = TRUE)
```
```{r}
dg %>% 
  count(country, year) %>% 
  arrange(desc(n))
```


```{r}
dg  %>% 
  filter(!is.na(SH.XPD.GHED.PP.CD)) %>% 
  group_by(country) %>% 
  summarise(mean=mean(SH.XPD.GHED.PP.CD, na.rm=TRUE), .groups="drop") %>% 
  arrange(desc(mean))
```






##  Current health expenditure per capita, PPP (current international $)

Series Code
:  SH.XPD.CHEX.PP.CD

```{r}
ce <- WDI(indicator = "SH.XPD.CHEX.PP.CD")
ce %>% head
```

```{r}
usethis::use_data(ce, overwrite = TRUE)

```
```{r}

ce %>% 
  count(country, year) %>% 
  arrange(desc(n))

```


## Adjusted net national income (annual % growth)

Series code
:  NY.ADJ.NNTY.KD.ZG
```{r}
adj <- WDI(indicator = "NY.ADJ.NNTY.KD.ZG")
adj %>% head
```

```{r}
usethis::use_data(adj, overwrite = TRUE)
```

```{r}
adj %>% 
  count(country, year) %>% 
  arrange(desc(n))
```
```{r}
d <- ce %>% 
  inner_join(adj, by = c('year','country')) %>% 
  inner_join(le, by = c('year','country')) %>% 
  select(-starts_with("iso")) %>% 
  select(country, year, everything())

d %>% head(10)
```
```{r}
usethis::use_data(d, overwrite = TRUE)
```




```{r}
country_choices <- 
  d %>% 
  select(country, SH.XPD.CHEX.PP.CD) %>% 
  na.omit() %>% 
  select(country) %>% 
  distinct() %>% 
  arrange() %>% 
  pull(country)
country_choices

```

```{r}
usethis::use_data(country_choices, overwrite = TRUE)
```
```{r}
series_choices <- d %>% 
select(-country, -year) %>% 
names
series_choices
```

```{r}
usethis::use_data(series_choices, overwrite = TRUE)
```

